---
title: "vegan_analysis"
format: html
editor: visual
---

```{r packages}
library("vegan")
```

```{r normal-fun}
mylog <- function(data) {
  if (any(data <= 0)) {
    return(log1p(data))
  } else {
    return(log(data))
  }
}

# normalization from grace et al 2016
# first does mylog (does natural logarithm. if 0s present, does log(one + x))
# then scales to z scores
# results it more normal looking data
normal <- function(data) {
  data %>%
    mylog() %>%
    scale() %>%
    as.numeric()
}
```

```{r data-prep}
bec_sample_loc <- here::here("data", "samples", "bec_extracted.shp")

if (!file.exists(bec_sample_loc)) {
  bec_samples_e <- extract(all_rasts, bec_samples, bind = T)
  writeVector(bec_samples_e, bec_sample_loc, overwrite = T)
}

bec_samples_e <- vect(bec_sample_loc)

bec_df <- bec_samples_e %>% 
  st_as_sf() %>%
  st_drop_geometry() %>%
  select(bec:VarDHI) %>%
  drop_na() %>%
  tibble() %>%
  filter(elev_cv <= 1000) %>%
  filter(elev_p95 >= 2000)

names_manual <- c(
  "bec",
  "category",
  "basal_area",
  "elev_cv",
  "elev_mean",
  "elev_p95",
  "elev_stddev",
  "gross_stem_volume",
  "loreys_height",
  "percentage_first_returns_above_2m",
  "percentage_first_returns_above_mean",
  "total_biomass",
  "CumDHI",
  "MinDHI",
  "VarDHI"
)

# names(bec_df) <- names(all_rasts)
names(bec_df) <- names_manual

bec_normal <- bec_df %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x))) 

bec_normal_ndup <- bec_df %>%
  group_by(across(c(bec, category, basal_area:total_biomass))) %>%
  dplyr::summarize(across(ends_with("DHI"), mean)) %>%
  ungroup() %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x))) 
```

```{r global-model}
to_rda <- bec_normal %>%
  select(-elev_mean,
         -percentage_first_returns_above_mean,
         -elev_stddev)

dhi_tab <- to_rda %>%
  select(ends_with("DHI"))

extracted <- to_rda %>%
  select(elev_p95,
         elev_cv,
         # elev_stddev,
         percentage_first_returns_above_2m)

modelled <- to_rda %>%
  select(loreys_height,
         basal_area,
         total_biomass,
         gross_stem_volume)



rda_out <- 
  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + loreys_height + total_biomass, # modelled
      data = to_rda)

rda_out

ordiplot(rda_out, choices = c(1, 2))

rda_eig <- rda_out$CCA$eig

# scores(rda_out, choices = c(1:3))$species
# scores(rda_out, choices = c(1:3))$biplot
# not actually the loadings, need to get those from correlations

summary(rda_out)$sites %>% 
  as.data.frame() %>% 
  select(RDA1) %>% 
  bind_cols(to_rda) %>%
  select(-bec, -category) %>% 
  cor() %>%
  as.data.frame() %>%
  select(RDA1) %>%
  arrange(desc(abs(RDA1)))

var_out <- varpart(dhi_tab, extracted, modelled)

varpart_out %>% 
  plot(Xnames = c("Derived", "Modelled"))

path_r2 <- var_out$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
  tibble()
```

Metrics derived solely from the point cloud explain 10.6% of the variation in the DHIs, while metrics modelled based on the point cloud explain 3.4% of the variation in the DHIs. 85.9% of the variation is unexplained by structural attributes.

```{r r2-global}
RsquareAdj(rda_out)$adj.r.squared
```

This model shows very little explanatory power. (13.9%)

```{r anova-global}
anova.cca(rda_out, step = 1000)

anova_out <- rda_out  %>%
  anova(by = "axis")

anova_out$`Pr(>F)` %>%
  as.data.frame() %>%
  rename(p_value = 1) %>%
  drop_na() %>%
  mutate(sig = cut(p_value, 
                   breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
                   labels = c("***", "**", "*", ".", "NS"))) %>%
  mutate(axis = paste0("RDA", row_number())) %>%
  relocate(axis)
```

The model is statistically significant (P \< 0.001). All of the axes are significant, although this may be related to the large sample size.

```{r bec-nest}
to_nest <- bec_normal

bec_nested <- to_nest %>%
  group_by(bec) %>%
  nest() %>%
  ungroup()
```

```{r rda-var-funs}
# function to take a tibble and get a list out outputs that can then be unnested and examined as needed. function(tibble) gets rda and varpart, as plottable objects ideally. then get the var_out$part$indfract. i'll also need to look at the significance using an anova at some point

my_rda_fun <- function(tibble) {
  dhi_tab <- tibble %>%
    select(ends_with("DHI"))
  
  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + loreys_height + total_biomass, # modelled, put condition here for a conditional rda
      data = tibble)
}

my_var_fun <- function(tibble) {
  dhi_tab <- tibble %>%
    select(ends_with("DHI"))
  
  extracted <- tibble %>%
    select(elev_p95,
           elev_cv,
           # elev_stddev,
           percentage_first_returns_above_2m)
  
  modelled <- tibble %>%
    select(loreys_height,
           basal_area,
           total_biomass,
           gross_stem_volume)
  
  varpart(dhi_tab, extracted, modelled)
}
```

```{r rda-var-bec}
outs <- bec_nested %>%
  mutate(rda = map(data, my_rda_fun),
         var = map(data, my_var_fun)) 
```

```{r fracts-out}

fracts_out <- outs %>%
  mutate(indfracts = map(var, .f = function(x) {
    x$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
      pivot_wider(names_from = type, values_from = Adj.R.squared)
    }
    )) %>%
  unnest(indfracts) %>%
  arrange(desc(residual)) %>%
  select(bec, extracted, modelled, overlap) %>%
  pivot_longer(c(extracted, modelled, overlap)) %>%
  mutate(value = ifelse(value > 0, value, 0)) %>%
  pivot_wider()
```

```{r raincloud-plot}
library(ggdist)
fracts_out %>%
  left_join(keys$bec, by = c("bec" = "zone")) %>%
  pivot_longer(extracted:overlap) %>%
  mutate(name = str_to_title(name)) %>%
  mutate(value = ifelse(value < 0, 0, value)) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye(
    # adjust bandwidth
    adjust = .5,
    # move to the right
    justification = -0.2,
    # remove the slub interval
    .width = 0,
    point_colour = NA
    ) +
  geom_boxplot(
    width = 0.12,
    # removing outliers
    outlier.color = NA,
    alpha = 0.5
  ) +
  stat_dots(
    # ploting on left side
    side = "left",
    # adjusting position
    justification = 1.1,
    # adjust grouping (binning) of observations
    binwidth = 0.0025,
    #position = "stack",
    aes(fill = group, col = group),
    alpha = 0.5
  ) +

  labs(y = NULL,
       x = "% Variance Explained",
       fill = "BEC Group",
       col = "BEC Group") +
  scale_x_continuous(labels = scales::label_percent()) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = c(0.8, .2))

```

```{r path-setup}
path_base <- tibble(x1 = c(0,
             0,
             0,
             0,
             1),
       y1 = c(1,
              -1,
              1,
              -1,
              0),
       x2 = c(1.85,
              1.85,
              .85,
              .85,
              1.85),
       y2 = c(.15,
              -.15,
              .15,
              -.15,
              0)) %>%
  ggplot(aes(x = x1, y = y1)) +
  geom_segment(aes(xend = x2, yend = y2), arrow = arrow()) +
  geom_text(aes(x = 0, y = -1.15), label = "Modelled") +
  geom_text(aes(x = 0, y = 1.15), label = "Extracted") +
  geom_text(aes(x = .65, y = 0), label = "Overlap") +
  geom_text(aes(x = 2, y = 0), label = "DHIs") +
  geom_point(data = tibble(x1 = c(.85, .85), 
                           y1 = c(-.5, .5)),
             size = 75,
             shape = 21,
             col = "grey") +
  theme_void() +
  scale_x_continuous(limits = c(-0.1, 2.1))

path_r2_plot <- tibble(type = c("extracted", "modelled", "overlap"),
       x1 = c(0.85, 0.85, 0.85),
       y1 = c(0.75, -0.75, 0))
```

```{r path-global}
var_oi <- c("VarDHI", "MinDHI", "CumDHI")
#var_oi <- c("VarDHI", "MinDHI")

var_out <- varpart(dhi_tab %>%
                     select(any_of(var_oi)),
                   extracted, modelled)

# plot(var_out, Xnames = c("Extracted", "Modelled"))

path_r2 <- var_out$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
  tibble()


path_r2_clean <- path_r2_plot %>%
  left_join(path_r2) %>%
  mutate(r2_label = paste0(round(Adj.R.squared * 100, 2), "%"))

path_base +
  geom_text(data = path_r2_clean, aes(label = r2_label)) +
  labs(title = paste(var_oi, collapse = " "))
```


```{r group-climate-spatial}
climate_folder <- here::here("data", "climate")

clim_vars <- climate_folder %>%
  list.files(pattern = "csv$") %>%
  tools::file_path_sans_ext() 

clim_vars <- clim_vars[!str_detect(clim_vars, "bec")]

group_clim_avg <-
  here::here(climate_folder, paste0(clim_vars, ".csv")) %>%
  read_csv() %>%
  mutate(var = clim_vars %>% rep(5) %>% sort) %>%
  select(group = BEC_CSV_GR, mean = MEAN, var)

```

```{r clim-plot-group}
fracts_clims <- fracts_out %>%
  left_join(keys$bec, by = c("bec" = "zone")) %>%
  group_by(group) %>%
  summarize(across(extracted:overlap, .fns = mean)) %>%
  left_join(group_clim_avg, multiple = "all")

clim_var = "tmax"

fracts_clims %>%
  pivot_longer(extracted:overlap) %>%
  filter(var == clim_var) %>%
  ggplot(aes(x = mean, y = value)) +
  geom_smooth(method = "lm", col = "red", lty = "dashed") +
  geom_point() +
  facet_wrap(~str_to_title(name)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x = clim_var,
       y = "% Variance Explained") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

```{r clim-plot-bec}
clim_vars <- climate_folder %>%
  list.files(pattern = "csv$") %>%
  tools::file_path_sans_ext() 

clim_vars <- clim_vars[!str_detect(clim_vars, "bec")]

bec_clim <- climate_folder %>%
  list.files(pattern = "bec.csv$",
             full.names = T) %>%
  read_csv() %>%
  mutate(var = clim_vars %>% rep(16) %>% sort) %>%
  select(bec = ZONE, mean = MEAN, var)

fracts_out %>%
  pivot_longer(extracted:overlap) %>%
  left_join(bec_clim, multiple = "all") %>%
  filter(var == "ppt") %>%
  ggplot(aes(x = mean, y = value)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(~name) 
```

```{r testing-idf-live}
to_rda <- bec_normal %>%
  filter(bec == "IDF")

dhi_tab <- to_rda %>%
  select(ends_with("DHI"))

extracted <- to_rda %>%
  select(elev_p95,
         elev_cv,
         percentage_first_returns_above_2m)

modelled <- to_rda %>%
  select(loreys_height,
         basal_area,
         total_biomass,
         gross_stem_volume)


rda_out <- 
  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + loreys_height + total_biomass, # modelled
    data = to_rda)

rda_out

ordiplot(rda_out, choices = c(1, 2))

rda_eig <- rda_out$CCA$eig

var_out <- varpart(dhi_tab, extracted, modelled)

plot(var_out, Xnames = c("Extracted", "Modelled"))

path_r2 <- var_out$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
  tibble()

path_r2_clean <- path_r2_plot %>%
  left_join(path_r2) %>%
  mutate(r2_label = paste0(round(Adj.R.squared * 100, 2), "%"))

path_base +
  geom_text(data = path_r2_clean, aes(label = r2_label)) +
  labs(title = "All")

summary(rda_out)$sites %>% 
  as.data.frame() %>% 
  select(RDA1) %>% 
  bind_cols(to_rda) %>%
  select(-bec, -category) %>% cor() %>%
  as.data.frame() %>%
  select(RDA1) %>%
  arrange(desc(abs(RDA1)))
```

```{r boxplots-dhi-bec}
bec_df %>%
  select(bec, ends_with("DHI")) %>%
  pivot_longer(cols = ends_with("DHI")) %>%
  group_by(name) %>% 
  left_join(keys$bec, by = c("bec" = "zone")) %>%
  mutate(bec = fct_reorder(bec, value)) %>%
  ggplot(aes(x = value, y = bec, fill = group)) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(x = "DHI Value",
       y = NULL,
       fill = NULL) +
  scale_fill_discrete_c4a_cat(palette = "batlow")
```





```{r cdf-anova}
outs %>% 
  filter(bec == "CDF") %>%
  select(rda) %>% 
  pull(rda) %>%
  .[[1]] %>%
  anova()
```

```{r cum-min-hexplot}
bec_df %>% 
  select(bec, CumDHI, MinDHI) %>%
  left_join(keys$bec, by = c("bec" = "zone")) %>%
  ggplot(aes(x = CumDHI, y = MinDHI)) +
  geom_hex() +
  facet_wrap(~bec)
```

```{r eig-anova}
outs %>%
  mutate(eig = map(rda, .f = function(rda) {
    rda$CCA$eig
  })) %>%
  unnest_wider(eig) %>%
  mutate(prop21 = RDA1/RDA2) %>%
  arrange(desc(prop21)) %>%
  left_join(anova_a2) %>%
  select(bec, RDA1:sig)

anova_outs <- outs %>%
  mutate(
    anova = map(rda, .f = function(rda) {
      anova_out <- rda  %>%
        anova(by = "axis")
      
      anova_out$`Pr(>F)` %>%
        as.data.frame() %>%
        rename(anova_p = 1) %>%
        drop_na() %>%
        mutate(anova_sig = cut(
          anova_p,
          breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
          labels = c("***", "**", "*", ".", "NS")))  %>%
        mutate(axis = paste0("RDA", row_number())) %>%
        relocate(axis)
  }))

anova_sig_axes <- anova_outs %>%
  select(bec, anova) %>%
  unnest(cols = anova) %>%
  filter(anova_sig != "NS")
```

```{r bec-loadings}
out_loadings <- outs %>%
  mutate(loadings = map(rda,.f = function(rda) {
    summary(rda)$sites %>%
      as.data.frame() %>%
      select(starts_with("RDA"))
  }),
  loadings = map2(.x = loadings, .y = data, .f = bind_cols),
  loadings = map(loadings, .f = function(x) {
    rcorr <- x %>% select(-category) %>%
      as.matrix() %>%
      Hmisc::rcorr()
    
    rvals <- rcorr$r %>%
      as.data.frame() %>%
      select(starts_with("RDA")) %>%
      mutate(var = row.names(.)) %>%
      pivot_longer(starts_with("RDA"),
                   values_to = "axis_loading",
                   names_to = "axis") %>%
      filter(!startsWith(var, "RDA"))
    
    pvals <- rcorr$P %>%
      as.data.frame() %>%
      select(starts_with("RDA")) %>%
      mutate(var = row.names(.)) %>%
      pivot_longer(starts_with("RDA"),
                   values_to = "axis_p",
                   names_to = "axis") %>%
      filter(!startsWith(var, "RDA"))
    
    left_join(rvals, pvals) %>%
      mutate(n = rcorr$n[[1, 1]])
  }
  
  )) %>%
  select(bec, loadings) %>%
  unnest(loadings)
```

```{r}
out_loadings %>% 
  mutate(axis_sig = cut(
          axis_p,
          breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
          labels = c("***", "**", "*", ".", "NS"),
          right = F)) %>%
  inner_join(anova_sig_axes) %>%
  #filter(axis_sig == "NS") %>%
  arrange(axis_loading) %>%
  count(var)

# need to remove the correlations for variables that arent actually involved
# percentage_first_returns_above_mean, elev_mean, elev_stddev
out_loadings %>% count(var)
```


need to get significant axes from the anova and then filter this

get total rda eigenvalues. find proportion of total for each rda. multiple by 1 - residual. and thats gets path number b/w RD1/RD2 and DHIs. can just filter each by which exist since only a few. goes for path plot

grab the p values w/in zones as well as the correlation coefficients using Hmisc::rcorr(). only plot these plus loadings over threshold

check if deciduous and coniferous have signfiicant axes